The Second Data Release of the INT Photometric H$\alpha$ Survey of the Northern Galactic Plane (IPHAS DR2)

### 2014-08-12 paper/iphas-dr2-rev2-arxiv.tex
Rev2 as uploaded to arxiv

% This is revision 2 of the IPHAS DR2 paper,
% created in response to the referee comments received on 11 August 2014
\documentclass[a4paper,useAMS,usenatbib]{mn2e}
\pdfpagewidth=\paperwidth
\pdfpageheight=\paperheight
\usepackage{amssymb,amsmath}
\usepackage[pdftex]{graphicx}
\usepackage{longtable} 
\usepackage{float}
\usepackage{booktabs}
\usepackage{aas_macros}
\usepackage{caption}
\usepackage{subcaption}

@@

et. al]{Geert Barentsen$^{1}$\thanks{E-mail:geert@barentsen.be},

@@

Having carried out object detection,
the astrometric calibration is then determined.
The solution starts with a rough 
World Coordinate System (WCS)
based on the known telescope and camera geometry,
using a Zenithal Polynomial projection
\citep[ZPN;][]{Calabretta2002}
to model the (fixed) field distortion of the camera.
The parameters of this solution
are then progressively refined by fitting against the
Two-Micron All Sky Survey \citep[2MASS; ][]{Skrutskie2006},
{ \bf
albeit without correcting for the $\sim$10 years of proper motion
between the IPHAS and 2MASS epochs. }
The resulting fit has previously been shown
to deliver results which are internally consistent
to better than 0\farcs 1 across the detector array
\citep{Gonzalez-Solares2008}.

An external validation of the astrometry 
has been carried out by comparing our positions against the
United States Naval Observatory CCD Astrograph Catalog 
\citep[UCAC4;][]{UCAC4}.
Fig.~\ref{fig:astrometry} shows the distribution
of the astrometric offsets
computed for 1.3 million stars
in the magnitude range $13 < r < 15$,
which is the range where both surveys overlap
and where the formal mean error of UCAC4
is better than 50~mas \citep{UCAC3}.
We find the mean difference in position 
between IPHAS and UCAC4 to be 94~mas,
which is satisfactory for our purposes.
{ \bf
The remaining residuals are in part due to
the motion of the Earth through our galaxy,
which we did not account for.
}

@@

(1) \emph{Depth.} 
We discarded any exposures 
for which the $5\sigma$ limiting magnitude\footnote{\bf
We defined the $5\sigma$ limiting magnitude
as the magnitude which a point source would have
if its flux equalled five times the level of the noise in the sky background.
The sky noise was estimated using a robust MAD estimator 
for noise scaled to equivalent Gaussian standard deviation,
ie. = MAD x 1.48,
after removing large scale sky background variations.
(MAD = Median of the Absolute Deviations about the median.)}
was brighter than 20th magnitude in the $r$-band
or brighter than 19th in $i$ or \ha.
Such data were typically obtained during poor weather or full moon.
Most observations were significantly better than these limits.
Fig.~\ref{fig:depth} presents the distribution of limiting magnitudes
for all quality-approved fields;
the mean depths and standard deviations are 
$21.2\pm0.5$ ($r$), $20.0\pm0.3$ ($i$) and $20.3\pm0.3$ (\ha).
The depth achieved depended
most strongly on the presence of the moon,
which was above the horizon during 62 per cent 
of the observations.
The great range in sky brightness this
produced is behind the wide and bi-modal shape
of the $r$-band limiting magnitude distribution 
(top panel in Fig.~\ref{fig:depth}).
In contrast, the depths attained in $i$ and \ha\ 
are less sensitive to moonlight, leading to
narrower magnitude limit distributions
(middle and bottom panels in Fig.~\ref{fig:depth}).
To a lesser extent, the wide spread in the $r$-band depth
is also explained by the shorter exposure time
that was used for this band during the first months of data-taking.

@@

\begin{table*}
\vspace{2.5cm}
\caption{\bf Recommended quality criteria 
for selecting objects with accurate colours.
These criteria serve to identify objects detected 
at S/N$\,>\,$10 in all three bands without being saturated,
and with the added requirement that the photometric measurements
need to be consistent across different aperture sizes.
The 86~million objects which satisfy these criteria
have been flagged in the catalogue using the column named \emph{a10}
(short for ``all-band 10$\sigma$'').}
\begin{tabular}{p{8cm}lp{6cm}}
  \hline
  Quality criterion & Rows passed & Description \\
  \hline
  $rErr\,<\,0.1$ {\sc and} $iErr\,<\,0.1$ {\sc and} $haErr\,<\,0.1$ &
  109 million (50\%) &
  Require the photon noise to be less than 
  0.1 mag in all bands (i.e. S/N$>$10).
  This implicitly requires a detection in all three bands.  \\
  {\sc not} \emph{saturated} &
  158 million (72\%) &
  The brightness must not exceed the nominal saturation limits.
  \\
  $|r- rAperMag1| 
  < 3\sqrt{rErr^2 + rAperMag1Err^2} 
  + 0.03$ &
  176 million (80\%) &
  Require the $r$ magnitude measured 
  in the default 2\farcs 3-diameter aperture
  to be consistent with the measurement 
  made in the smaller 1\farcs 2 aperture,
  albeit tolerating a 0.03 mag systematic error.
  This will reject sources for which the background
  subtraction or the deblending procedure
  was not performed reliably. \\
  $|i- iAperMag1| 
  < 3\sqrt{iErr^2 + iAperMag1Err^2} 
  + 0.03$ &
  183 million (84\%) &
  Same as above for $i$. \\
  $|ha- haAperMag1| < 
  3\sqrt{haErr^2 + haAperMag1Err^2} 
  + 0.03$ &
  158 million (72\%) &
  Same as above for \ha. \\
  \hline
  All of the above (flagged as {\bf\emph{a10}}) &
  86 million (39\%) & \\
  \hline
\end{tabular}
\label{tab:a10}
\vspace{2.5cm}
\caption{\bf Additional quality criteria which are recommended
for applications which require objects to be
single, unconfused point sources with accurate colours.
The 59 million sources which satisfy these criteria 
have been flagged using the column named \emph{a10point}.}
\begin{tabular}{p{8cm}lp{6cm}}
  \hline
  Quality criterion & Rows passed & Description \\
  \hline
   \emph{a10} &
   86 million (39\%) &
   The object must satisfy the criteria for accurate colours listed in Table~\ref{tab:a10}. \\

   $pStar > 0.9$ &
   145 million (66\%) &
   % pStar > 0.89 is identical to mergedClass == -1
   The object must appear as a perfect point source,
   as inferred from comparing its PSF
   with the average PSF measured in the same CCD. \\

@@

@@

\begin{figure}
    \includegraphics[width=0.5\textwidth]{figures/magdist/magdist-r.pdf} 
    \caption{r-band magnitude distribution
    for all objects in the catalogue (light grey).
    Overlaid we also show the distribution for objects
    detected at S/N $>$ 10 in all bands
    selected according to the quality criteria 
    described in Table~\ref{tab:a10} (\emph{a10}, grey),
    and for the set of unconfused 10$\sigma$ point source detections
    described in Table~\ref{tab:a10point} (\emph{a10point}, dark grey).
    The distributions for $i$ and \ha\
    look identical, apart from being shifted
    by about 1 and 0.5 mag towards brighter magnitudes,
    respectively.}
    \label{fig:magdist}
\end{figure}

@@

    \vspace{2cm}
    \includegraphics[width=\textwidth]{figures/sourcecount/sourcecount.pdf} 
    \caption{Number of entries in the IPHAS DR2 source catalogue
    as a function of Galactic longitude.
    The upper blue line shows the number of sources
    counted in 1\degr-wide longitude bins.
    The lower red line uses the same binning
    but includes only those sources 
    for which the \emph{deblend} flag is {\sc false}, 
    i.e. unconfused sources for which the CASU pipeline
    did not have to apply a deblending procedure.
    In both cases we counted only those sources
    in the latitude range $-5\degr<b<+5\degr$,
    such that one may obtain a rough guide to source density
    by dividing the counts by 10~deg$^2$.
    The global variations in the source counts
    traces the structure of the Galaxy
    and the distribution of foreground extinction,
    but is also affected by instrumental effects
	such as variations in the observed depth
	and completeness (see text).
   }
    \label{fig:sourcecount}
\end{figure*}

@@

>>>>

\documentclass[a4paper,useAMS,usenatbib]{mn2e}
\pdfpagewidth=\paperwidth
\pdfpageheight=\paperheight
\usepackage{amssymb,amsmath}
\usepackage[pdftex]{graphicx}
\usepackage{longtable} 
\usepackage{float}
\usepackage{booktabs}
\usepackage{aas_macros}
\usepackage{caption}
\usepackage{subcaption}

@@

et al.]{Geert Barentsen$^{1}$\thanks{E-mail:geert@barentsen.be},

@@

Having carried out object detection,
the astrometric calibration is then determined.
The solution starts with a rough 
World Coordinate System (WCS)
based on the known telescope and camera geometry,
using a Zenithal Polynomial projection
\citep[ZPN;][]{Calabretta2002}
to model the (fixed) field distortion of the camera.
The parameters of this solution
are then progressively refined by fitting against the
Two-Micron All Sky Survey \citep[2MASS; ][]{Skrutskie2006},
albeit without correcting for the $\sim$10 years of proper motion
between the IPHAS and 2MASS epochs.
The resulting fit has previously been shown
to deliver results which are internally consistent
to better than 0\farcs 1 across the detector array
\citep{Gonzalez-Solares2008}.

An external validation of the astrometry 
has been carried out by comparing our positions against the
United States Naval Observatory CCD Astrograph Catalog 
\citep[UCAC4;][]{UCAC4}.
Fig.~\ref{fig:astrometry} shows the distribution
of the astrometric offsets
computed for 1.3 million stars
in the magnitude range $13 < r < 15$,
which is the range where both surveys overlap
and where the formal mean error of UCAC4
is better than 50~mas \citep{UCAC3}.
We find the mean difference in position 
between IPHAS and UCAC4 to be 94~mas,
which is satisfactory for our purposes.
These residuals are in part due to
the motion of the Earth through our Galaxy,
which we did not account for.

@@

(1) \emph{Depth.} 
We discarded any exposures 
for which the $5\sigma$ limiting magnitude\footnote{
We defined the $5\sigma$ limiting magnitude
as the magnitude which a point source would have
if its flux equalled five times the level of the noise in the sky background.
The sky noise was estimated using a robust MAD estimator 
for noise scaled to equivalent Gaussian standard deviation,
ie. = MAD x 1.48,
after removing large scale sky background variations.
(MAD = Median of the Absolute Deviations about the median.)}
was brighter than 20th magnitude in the $r$-band
or brighter than 19th in $i$ or \ha.
Such data were typically obtained during poor weather or full moon.
Most observations were significantly better than these limits.
Fig.~\ref{fig:depth} presents the distribution of limiting magnitudes
for all quality-approved fields;
the mean depths and standard deviations are 
$21.2\pm0.5$ ($r$), $20.0\pm0.3$ ($i$) and $20.3\pm0.3$ (\ha).
The depth achieved depended
most strongly on the presence of the moon,
which was above the horizon during 62 per cent 
of the observations.
The great range in sky brightness this
produced is behind the wide and bi-modal shape
of the $r$-band limiting magnitude distribution 
(top panel in Fig.~\ref{fig:depth}).
In contrast, the depths attained in $i$ and \ha\ 
are less sensitive to moonlight, leading to
narrower magnitude limit distributions
(middle and bottom panels in Fig.~\ref{fig:depth}).
To a lesser extent, the wide spread in the $r$-band depth
is also explained by the shorter exposure time
that was used for this band during the first months of data-taking.

@@

\begin{table*}
\vspace{2.5cm}
\caption{Recommended quality criteria 
for selecting objects with accurate colours.
These criteria serve to identify objects detected 
at S/N$\,>\,$10 in all three bands without being saturated,
and with the added requirement that the photometric measurements
need to be consistent across different aperture sizes.
The 86~million objects which satisfy these criteria
have been flagged in the catalogue using the column named \emph{a10}
(short for ``all-band 10$\sigma$'').}
\begin{tabular}{p{8cm}lp{6cm}}
  \hline
  Quality criterion & Rows passed & Description \\
  \hline
  $rErr\,<\,0.1$ {\sc and} $iErr\,<\,0.1$ {\sc and} $haErr\,<\,0.1$ &
  109 million (50\%) &
  Require the photon noise to be less than 
  0.1 mag in all bands (i.e. S/N$>$10).
  This implicitly requires a detection in all three bands.  \\
  {\sc not} \emph{saturated} &
  158 million (72\%) &
  The brightness must not exceed the nominal saturation limits.
  \\
  $|r- rAperMag1| 
  < 3\sqrt{rErr^2 + rAperMag1Err^2} 
  + 0.03$ &
  176 million (80\%) &
  Require the $r$ magnitude measured 
  in the default 2\farcs 3-diameter aperture
  to be consistent with the measurement 
  made in the smaller 1\farcs 2 aperture,
  albeit tolerating a 0.03 mag systematic error.
  This will reject sources for which the background
  subtraction or the deblending procedure
  was not performed reliably. \\
  $|i- iAperMag1| 
  < 3\sqrt{iErr^2 + iAperMag1Err^2} 
  + 0.03$ &
  183 million (84\%) &
  Same as above for $i$. \\
  $|ha- haAperMag1| < 
  3\sqrt{haErr^2 + haAperMag1Err^2} 
  + 0.03$ &
  158 million (72\%) &
  Same as above for \ha. \\
  \hline
  All of the above (flagged as {\bf\emph{a10}}) &
  86 million (39\%) & \\
  \hline
\end{tabular}
\label{tab:a10}
\vspace{2.5cm}
\caption{Additional quality criteria which are recommended
for applications which require objects to be
single, unconfused point sources with accurate colours.
The 59 million sources which satisfy these criteria 
have been flagged using the column named \emph{a10point}.}
\begin{tabular}{p{8cm}lp{6cm}}
  \hline
  Quality criterion & Rows passed & Description \\
  \hline
   \emph{a10} &
   86 million (39\%) &
   The object must satisfy the criteria for accurate colours listed in Table~\ref{tab:a10}. \\

   $pStar > 0.9$ &
   145 million (66\%) &
   The object must appear as a perfect point source,
   as inferred from comparing its PSF
   with the average PSF measured in the same CCD. \\

@@

\begin{figure}
    \includegraphics[width=0.45\textwidth]{figures/magdist/magdist-r.pdf} 
    \caption{r-band magnitude distribution
    for all objects in the catalogue (light grey).
    Overlaid we also show the distribution for objects
    detected at S/N $>$ 10 in all bands
    selected according to the quality criteria 
    described in Table~\ref{tab:a10} (\emph{a10}, grey),
    and for the set of unconfused 10$\sigma$ point source detections
    described in Table~\ref{tab:a10point} (\emph{a10point}, dark grey).
    The distributions for $i$ and \ha\
    look identical, apart from being shifted
    by about 1 and 0.5 mag towards brighter magnitudes,
    respectively.}
    \label{fig:magdist}
\end{figure}

@@

@@

    \vspace{1.5cm}
    \includegraphics[width=\textwidth]{figures/sourcecount/sourcecount.pdf} 
    \caption{Number of entries in the IPHAS DR2 source catalogue
    as a function of Galactic longitude.
    The upper blue line shows the number of sources
    counted in 1\degr-wide longitude bins.
    The lower red line uses the same binning
    but includes only those sources 
    for which the \emph{deblend} flag is {\sc false}, 
    i.e. unconfused sources for which the CASU pipeline
    did not have to apply a deblending procedure.
    In both cases we counted only those sources
    in the latitude range $-5\degr<b<+5\degr$,
    such that one may obtain a rough guide to source density
    by dividing the counts by 10~deg$^2$.
    The global variations in the source counts
    traces the structure of the Galaxy
    and the distribution of foreground extinction,
    but is also affected by instrumental effects
	such as variations in the observed depth
	and completeness (see text).
   }
    \label{fig:sourcecount}
\end{figure*}

@@

% woof woof ／人 ◕‿‿◕ 人＼ woof woof
